---
title: "loops_and_functions"
author: "Charles Hendrickson"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Task 1 
Write a nested for loop that creates a matrix containing the following:

```{r}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

transects <- c("Transect A", "Transect B", "Transect C")

# Need to create a MATRIX to store the outputs in: 
out_matrix <- matrix(nrow = length(weekdays), ncol = length(transects))


# Now, a nested for loop:
for (i in seq_along(weekdays)) { # Outer loop is the weekdays
  for (j in seq_along(transects)) { # Inner loop is the transects
    out_val <- print(paste(weekdays[i], "-", transects[j]))
    out_matrix[i] <- out_val
  }
}

```

#Task 2
Write a function called force that calculates a force (in Newtons), given inputs of mass (in kg) and acceleration (in ms2 (recall: F=ma), and returns a statement “The resulting force is ___ Newtons.”

```{r}
force <- function(mass_kg, acceleration_ms2) {
  print(paste("The resulting force is", mass_kg * acceleration_ms2, "Newtons"))
}

force(mass_kg = 2, acceleration_ms2 = 2)
```

#Task 3
Write a function called name_column that, given a data frame name and a number, will return the column name associated with that column number. Add an error message if the column number is < 1, or if the number exceeds the number of columns in the data frame. Test the function using the mtcars data frame.

```{r}
name_column <- function(df_name, num) {
    if (num >= 1) {
    return(colnames(df_name[num]))
} else if (num < 1 | num > ncol(df_name) ) {
  stop("Column number must be greater than 1 or the number exceeds the number of columns in the data frame.")}
}

name_column(mtcars, 1)
```

# Task 3 Part 2
W=aLb, where where L is total fish length (centimeters), W is the expected fish weight (grams), and a and b are species-dependent parameter values 

Recreate the table above as a data frame stored as fish_parms. Then, write a function called fish_weight that allows a user to only enter the common name (argument fish_name) and total length (argument tot_length) (in centimeters) of a fish, to return the expected fish weight in grams. Test it out for different species and lengths.

Now, try creating a vector of lengths (e.g. 0 to 100, by increments of 1) and ensuring that your function will calculate the fish weight over a range of lengths for the given species (try this for milkfish, storing the output weights as milkfish_weights.
```{r}
# My function
fish_weight <- function(fish_name, length_cm) {

# Data frame:
fish_parms <- data.frame(sci_name = c("Chanos chanos", "Sphyraena barracuda", "Caranx ignobilis"), 
                         common_name = c("Milkfish", "Great barracuda", "Giant trevally"), 
                         a_est = c(0.0905, 0.0181, 0.0353),
                         b_est = c(2.52, 3.27, 3.05))


# Telling R which "common_name" to look for 
fish_choice <- dplyr::filter(.data = fish_parms, common_name == fish_name)


# Value that my function will return
weight_g <- fish_choice$a_est * length_cm ^ fish_choice$b_est

# What you want to return back to the user
return(weight_g)
}


fish_weight("Great barracuda", 3)
```

# Task 4

```{r}
wave_power <- function(height, period) {
 power <- return((0.5 * height ^ 2)* period)
}

seq_wave_height <- seq(from = 0,to = 3, by = 0.2)

wave_power(seq_wave_height, 8)

```

# Task #5
```{r}
final_wave_power <- function(height, period, water_depth) {
  if (water_depth > 12) {
    message("Depths greater than 12m use shallow_wave_power.")
      power <- (0.5 * height ^ 2)* period
    
} else if (water_depth <= 12) { 
      power <- (0.81 * height ^ 2)* period
      message("Depths less than or equal to 12m use shallow_wave_power.")
  }
  return(power)
}

final_wave_power(5, 2, 8)
```


































